The Elements of Theses

### 2014-09-08 thesis/DataManagement.tex
update to thesis
fix typos

The rest of this chapter is organized as follows.
Section~\ref{data:options} reviews the background and discusses the design options for snapshot backup 
with a VM-centric approach. 
Section~\ref{data:arch} describes our system architecture and implementation details.
Section~\ref{data:analysis} analyzes the tradeoff and benefits of our approach. 
Section~\ref{data:deletion} describes out approximate deletion algorithm.
%   the benefit of our approach for fault isolation. 
Section~\ref{data:impl} is our experimental evaluation that compares with other approaches.
Section~\ref{data:related} reviews the related works.
Section~\ref{data:concl} concludes this chapter.

@@

Additional APIs such as \textbf{Scan()}, \textbf{Compact()}, \textbf{Create()} and \textit{Remove()} are provided for 
higher level operations. Scan() is used by the map-reduce procedure to collect the most popular blocks among all snapshot stores.
Compact() is called when the system determines a container has too much unclaimed space and needs to be reclaimed.
Create() and Remove() initialize and delete a container respectively.

\section{Analysis of VM-centric Approach}
\label{data:analysis}
In this section we give out analysis of the impacts of our VM-centric deduplication scheme. 
As introduced in chapter~\ref{inline:dedup}, our VM-centric approach put the cluster wide 
popular data into PDS which prevent each VM from keeping their own copies. We will show that by adding
extra replication to PDS data, this strategy brings advantages to both fault tolerancy and deduplication efficiency.
The parameters we will use in our analysis below are defined in Table~\ref{tab:symbol}.

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used in the VM snapshot recipes.
A mark-and-sweep process  requires a scan of the entire snapshot store.
%Our VM-centric mark-and-sweep procedure is similar to Guo\cite{Guo2011} 
%in a way that the data dependency is clearly known.
%This generally requires a scan of almost the entire snapshot store which is much more expensive than approximate deletion,
%although certain optimized algorithms may apply\cite{Przemyslaw2013}. But 
Since it is a VM-specific procedure, 
the space and time cost is relatively small compared to the tranditional mark-and-sweep
which scans snapshot chunks from all VMs.
For example, consider each reference consumes 8 bytes plus  1 mark bit. A VM that has 40GB backup data with about
10 million chunks will need less than 85MB of memory to complete a VM-specific mark-and-sweep process
in less than half an hour, assuming 50MB/s disk bandwidth is allocated.
\end{itemize}

@@

Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC,
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
namely the percent of duplicate chunks
which are detected and removed.
% so only global perfect deduplication can have 100\% efficiency.
% We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB per machine
and  the deduplication efficiency can reach over 96.33\%.
When $\sigma=4\%$, the deduplication efficiency can reach 96.9\% while space consumption increases to 200MB per machine. 
%and  the deduplication efficiency can reach 96.489\%.
%When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster for fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB can deliver up to 97.79\% deduplication efficiency, which is slightly better than VC.
Thus this represents a tradeoff that VC provides better fault tolerance and fast approximate deletion
with competitive deduplication efficiency.

% SRB can deliver up to 97.935\% deduplication efficiency which is slightly better than VC.
% Thus this represents a tradeoff: VC provides better fault tolerance and fast
% approximate deletion with competitive (but slightly lower) deduplication
% efficiency.
% >>>>>>> 4dcd68612b052c0f890ed30361a9ed03771b5df1
%its loss of efficiency is caused by the routing of data chunks which restricts the search scope
%of duplicate detection.
%SRB provides slightly better deduplication efficiency than our VC approach, which we accept
%because we are rewarded by architecture-wise benefits such as fault-tolerance and fast 
%Though not our primary goal, this result does show that VC can remove competitive amount of duplicates.

@@

\section{Related Work}
\label{data:related}
Since it is expensive to compare a large number of chunk signatures for deduplication,
several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR,arrow08}.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements with the tradeoff of a reduced deduplication ratio.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault tolerance in the cluster-based environment that we have considered
in this work.

In designing a VC duplication solution, we have considered and adopted some of
the following previously-developed techniques.
1)
{\em Changed block Tracking}.
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011,cvfs03}.
Such a scheme  is  VM-centric since deduplication is localized. 
We are seeking for a tradeoff since 
global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
2) {\em Stateless Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing by Dong et al.~\cite{Dong2011} 
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index}.
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

\section{Concluding Remarks}
\label{data:concl}
In this chapter we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The key contribution is a VM-centric data management architecture that
integrate with multi-level deduplication to 
maximize fault isolation while delivering competitive deduplication efficiency.
Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
while global deduplication with a small common data set eliminates popular duplicates.
VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that VC strikes a tradeoff and 
can accomplish 96.33\% or 96.9\% of what complete global
deduplication can do.  The availability of snapshots increases substantially with 
a small replication overhead for popular inter-VM chunks.

>>>>

The rest of this chapter is organized as follows.
Section~\ref{data:options} reviews the background and discusses the design options for snapshot backup 
with a VM-centric approach. 
Section~\ref{data:arch} describes our system architecture and implementation details.
Section~\ref{data:analysis} analyzes the trade-off and benefits of our approach.
Section~\ref{data:deletion} describes out approximate deletion algorithm.
%   the benefit of our approach for fault isolation. 
Section~\ref{data:impl} is our experimental evaluation that compares with other approaches.
Section~\ref{data:related} reviews the related works.
Section~\ref{data:concl} concludes this chapter.

@@

Additional APIs such as \textbf{Scan()}, \textbf{Compact()}, \textbf{Create()} and \textbf{Remove()} are provided for 
higher level operations. Scan() is used by the map-reduce procedure to collect the most popular blocks among all snapshot stores.
Compact() is called when the system determines a container has too much unclaimed space and needs to be reclaimed.
Create() and Remove() initialize and delete a container respectively.

\section{Analysis of VM-centric Approach}
\label{data:analysis}
In this section we give out analysis of the impacts of our VM-centric deduplication scheme. 
As introduced in chapter~\ref{inline:dedup}, our VM-centric approach put the cluster wide 
popular data into PDS which prevent each VM from keeping their own copies. We will show that by adding
extra replication to PDS data, this strategy brings advantages to both fault tolerance and deduplication efficiency.
The parameters we will use in our analysis below are defined in Table~\ref{tab:symbol}.

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used in the VM snapshot recipes.
A mark-and-sweep process  requires a scan of the entire snapshot store.
%Our VM-centric mark-and-sweep procedure is similar to Guo\cite{Guo2011}
%in a way that the data dependency is clearly known.
%This generally requires a scan of almost the entire snapshot store which is much more expensive than approximate deletion,
%although certain optimized algorithms may apply\cite{Przemyslaw2013}. But
Since it is a VM-specific procedure,
the space and time cost is relatively small compared to the transitional mark-and-sweep
which scans snapshot chunks from all VMs.
For example, consider each reference consumes 8 bytes plus  1 mark bit. A VM that has 40GB backup data with about
10 million chunks will need less than 85MB of memory to complete a VM-specific mark-and-sweep process
in less than half an hour, assuming 50MB/s disk bandwidth is allocated.
\end{itemize}

@@

Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC,
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
namely the percent of duplicate chunks
which are detected and removed.
% so only global perfect deduplication can have 100\% efficiency.
% We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB per machine
and  the deduplication efficiency can reach over 96.33\%.
When $\sigma=4\%$, the deduplication efficiency can reach 96.9\% while space consumption increases to 200MB per machine. 
%and  the deduplication efficiency can reach 96.489\%.
%When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster for fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB can deliver up to 97.79\% deduplication efficiency, which is slightly better than VC.
Thus this represents a trade-off that VC provides better fault tolerance and fast approximate deletion
with competitive deduplication efficiency.

% SRB can deliver up to 97.935\% deduplication efficiency which is slightly better than VC.
% Thus this represents a trade-off: VC provides better fault tolerance and fast
% approximate deletion with competitive (but slightly lower) deduplication
% efficiency.
% >>>>>>> 4dcd68612b052c0f890ed30361a9ed03771b5df1
%its loss of efficiency is caused by the routing of data chunks which restricts the search scope
%of duplicate detection.
%SRB provides slightly better deduplication efficiency than our VC approach, which we accept
%because we are rewarded by architecture-wise benefits such as fault-tolerance and fast
%Though not our primary goal, this result does show that VC can remove competitive amount of duplicates.

@@

\section{Related Work}
\label{data:related}
Since it is expensive to compare a large number of chunk signatures for deduplication,
several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08}
uses  an in-memory Bloom filter and a prefetching cache for data chunks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR,arrow08}.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}
to reduce memory requirements with the trade-off of a reduced deduplication ratio.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is
studied in ~\cite{wei2013}.
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault tolerance in the cluster-based environment that we have considered
in this work.

In designing a VC duplication solution, we have considered and adopted some of
the following previously-developed techniques.
1)
{\em Changed block Tracking}.
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011,cvfs03}.
Such a scheme  is  VM-centric since deduplication is localized.
We are seeking for a trade-off since
global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
2) {\em Stateless Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing by Dong et al.~\cite{Dong2011}
Stateless data routing divides the deduplication work with a similarity approximation. This work
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index}.
One effective approach that reduces memory usage is
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}.
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture.
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

\section{Concluding Remarks}
\label{data:concl}
In this chapter we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The key contribution is a VM-centric data management architecture that
integrate with multi-level deduplication to
maximize fault isolation while delivering competitive deduplication efficiency.
Similarity guided local search reduces cross-VM data dependency and exposes more parallelism
while global deduplication with a small common data set eliminates popular duplicates.
VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that VC strikes a trade-off and
can accomplish 96.33\% or 96.9\% of what complete global
deduplication can do.  The availability of snapshots increases substantially with 
a small replication overhead for popular inter-VM chunks.

### 2014-09-08 thesis/OfflineDedup.tex
fix typos

While our synchronous offline solution accomplishes perfect deduplication efficiency while maintaining
low resource usage,
the tradeoff of this approach is that the job turnaround time for each individual
VM backup task becomes longer. In addition,
every machine has to read dirty segments twice
and that some deduplication requests are delayed for staged parallel processing.
However, with our careful parallelism and buffer management,
this multi-stage detection scheme can provide a sufficient throughput for VM backup.

@@

%While we assume all machines can process  independent or coordinated
%deduplication and backup operations in parallel,  
%our scheme uses a single-thread procedure on each machine to minimize the use of CPU, network bandwidth, and disk bandwidth.
We use the two level metadata hierachy as discussed in section~\ref{overview:meta}.
The snapshot metadata  contains a list of segments and other meta data information.
Segment metadata contains its  content block fingerprints and reference pointers. 
If a segment is not changed from one snapshot to another, indicated by a dirty bit embedded in the virtual disk driver, 
%its content blocks are not changed as well, thus 
its segment metadata contains a reference pointer to an earlier segment.
For a dirty segment, if one of its blocks is duplicate to another block in the system,  
the block metadata contains a reference pointer to the earlier block.

@@

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is 
used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
As discussed in Section~\ref{offline:intro},
there is no dedicated resource for deduplication in our targeted setting and low memory usage
 is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.

@@

\section{Concluding Remarks}
\label{offline:concl}
The contribution  of this work is a low-cost multi-stage parallel deduplication solution.
Because of separation  of duplicate detection and actual backup,
we are able to evenly distribute  fingerprint comparison among clustered machine
nodes, and only load one partition at time at each machine for in-memory comparison.
%The tradeoff is that every machine has to read dirty segments twice 
%and that some deduplication requests are delayed for staged processing.

>>>>

While our synchronous offline solution accomplishes perfect deduplication efficiency while maintaining
low resource usage,
the trade-off of this approach is that the job turnaround time for each individual
VM backup task becomes longer. In addition,
every machine has to read dirty segments twice
and that some deduplication requests are delayed for staged parallel processing.
However, with our careful parallelism and buffer management,
this multi-stage detection scheme can provide a sufficient throughput for VM backup.

@@

%While we assume all machines can process  independent or coordinated
%deduplication and backup operations in parallel,
%our scheme uses a single-thread procedure on each machine to minimize the use of CPU, network bandwidth, and disk bandwidth.
We use the two level metadata hierarchy as discussed in section~\ref{overview:meta}.
The snapshot metadata  contains a list of segments and other meta data information.
Segment metadata contains its  content block fingerprints and reference pointers.
If a segment is not changed from one snapshot to another, indicated by a dirty bit embedded in the virtual disk driver,
%its content blocks are not changed as well, thus
its segment metadata contains a reference pointer to an earlier segment.
For a dirty segment, if one of its blocks is duplicate to another block in the system,
the block metadata contains a reference pointer to the earlier block.

@@

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is
used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08}
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
As discussed in Section~\ref{offline:intro},
there is no dedicated resource for deduplication in our targeted setting and low memory usage
 is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011}
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}
to reduce memory requirement with a trade-off of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup
In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.

@@

\section{Concluding Remarks}
\label{offline:concl}
The contribution  of this work is a low-cost multi-stage parallel deduplication solution.
Because of separation  of duplicate detection and actual backup,
we are able to evenly distribute  fingerprint comparison among clustered machine
nodes, and only load one partition at time at each machine for in-memory comparison.
%The trade-off is that every machine has to read dirty segments twice
%and that some deduplication requests are delayed for staged processing.
